Optimal Symmetric Multimodal Templates and Concatenated Random Forests for Supervised Brain Tumor Segmentation (Simplified) with \textit{ANTsR}

### 2014-04-28 paper.tex
WIP: minor comments

In line with our philosophy regarding open science \citep{tustison2013,ince2012}, 
we have made the code available as open source%
\footnote{
https://github.com/ntustison/BRATS2013
}
and, to our knowledge, we are the only competitors out of the approximately 20 teams 
that participated in either years' challenges to do this.  All code is based
on the well-vetted and open source Insight Toolkit (ITK)%
\footnote{
http://www.itk.org/
} 
of the National Institutes of Health and is available through our \textit{ANTsR} 
package.
In subsequent sections we describe in greater detail the automated 
supervised brain segmentation pipeline including the generation of asymmetry
feature images and the concatenated RF construction.  We also
describe the code and data%
%\footnote{
%Although all data was provided having been skull-stripped, 
%we note to the reader that an
%available ANTs-based solution to the brain extraction problem
%is found in the script {\tt antsBrainExtraction.sh} which uses
%a hybrid approach based on image registration and binary morphology
%to isolate the brain \citep{avants2010a}.  An evaluation with other
%methodologies demonstrated comparable to superior performance 
%\citep{avants2010a}.
%} 
used for the BRATS challenge so that 
the interested reader can reproduce our results.  Finally, we 
illustrate performance with results from the BRATS 2013 challenge.

@@

Our segmentation protocol involves training and application
of two RF models in succession 
(see Figure \ref{fig:pipeline}) to which we refer as ``stages.''  The basic idea is that
we generate a set of feature images used as input to the
first RF model (or first stage) which produces a voxelwise 
probabilistic tissue estimate%
\footnote{
More precisely, in the RF framework, each prediction sample (i.e. the feature vector at each voxel),
is propagated through each tree of the ensemble where it is labeled as belonging to a specific
class.  These ``votes'' are converted to voxelwise probabilistic estimates for each class.
}
 for the seven brain/tumor tissue types:
\begin{itemize}
\item cerebrospinal fluid,
\item grey matter,
\item white matter,
\item edema, 
\item non-enhancing tumor (including low-grade tumor center), 
\item enhancing tumor (excluding necrotic center), and 
\item abnormal necrotic center or necrocyst in high-grade gliomas.
\end{itemize}
We then use these output tissue map estimates as spatial 
priors for generating a second set of image features.  These
are used as input for a second RF model application (or
second stage),
the output of which constitute the final tumor segmentation estimate.
Note that some feature images are included in both stages such as
the asymmetry features described in the previous section.

>>>>

In line with our philosophy regarding open science \citep{tustison2013,ince2012}, 
we have made the code available as open source%
\footnote{
https://github.com/ntustison/BRATS2013
}
and, to our knowledge, we are the only competitors out of the approximately 20 teams 
that participated in either years' challenges to do this.  All code is based
on the well-vetted and open source Insight Toolkit (ITK)%
\footnote{
http://www.itk.org/
} 
of the National Institutes of Health and is available through our \textit{ANTsR} 
package.
In subsequent sections we describe in greater detail the automated 
supervised brain segmentation pipeline including the generation of asymmetry
feature images and the concatenated RF construction.  We also
describe the code and data used for the BRATS challenge so that 
the interested reader can reproduce our results.  Finally, we 
illustrate performance with results from the BRATS 2013 challenge.

%\footnote{
%Although all data was provided having been skull-stripped, 
%we note to the reader that an
%available ANTs-based solution to the brain extraction problem
%is found in the script {\tt antsBrainExtraction.sh} which uses
%a hybrid approach based on image registration and binary morphology
%to isolate the brain \citep{avants2010a}.  An evaluation with other
%methodologies demonstrated comparable to superior performance 
%\citep{avants2010a}.
%}

@@

Our segmentation protocol involves training and application
of two RF models in succession 
(see Figure \ref{fig:pipeline}) to which we refer as ``stages.''  The basic idea is that
we generate a set of feature images used as input to the
first RF model (or first stage) which produces a voxelwise 
probabilistic tissue estimate.  More precisely, in the RF framework, each prediction sample (i.e. the feature vector at each voxel),
is propagated through each tree of the ensemble where it is labeled as belonging to a specific
class.  These ``votes'' are converted to voxelwise probabilistic
estimates for each class.
\textcolor{red}{FIXME --- above comments need more detail ... how/why
  is this a probability, what does this mean?   it's bootstrapping etc
right so what can we say about this?}
The seven brain/tumor tissue types include:
\begin{itemize}
\item cerebrospinal fluid,
\item grey matter,
\item white matter,
\item edema, 
\item non-enhancing tumor (including low-grade tumor center), 
\item enhancing tumor (excluding necrotic center), and 
\item abnormal necrotic center or necrocyst in high-grade gliomas.
\end{itemize}
We then use these output tissue map estimates as spatial 
priors for generating a second set of image features.  These
are used as input for a second RF model application (or
second stage),
the output of which constitute the final tumor segmentation estimate.
Note that some feature images are included in both stages such as
the asymmetry features described in the previous section.
